\chapter{Notes}
\label{cha:notes}

\section{Experimental Setup/Analysis}
Processing/Algorithm is well known, reliance on pure hardware-events input.
------------

\section{Related Research}
 Related Research: 
 ------------------------
 As a noticeable result, Fan et al. [5]
 have proposed a corrected method to construct the statistical
 value for runs distribution test. They have shown that there
 exist some inaccuracies in the runs distribution test of NIST
 SP 800-22, and the degrees of freedom of the test statistic
 is adjusted. 
  \cite{kangadditional}
------------------------------
 Sehr aehnlicher ansatz, tracken von cycles, allerings sehr viel einfachere Bewertung 
 \cite{yoo2017analysis}

\section{NIST}


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End: 

% NIST
---------------------------------------------------------
% https://github.com/usnistgov/SP800-90B_EntropyAssessment
 Number od reuiqred Samples 19 / 11 \cite{turan2018nist} 
The submitter makes an IID claim on the noise source, based on the submitter?s analysis of the design. The submitter shall provide rationale for the IID claim. \cite{turan2018nist} -> Auf MD5 verweisen
% SP 800-90B Introduction
 The test suite of SP 800-90B is divided into two major steps. The first step is to determine the track, IID(independent and identically distributed) or Non-IID, and the second step is to estimate the entropy of the given source. The permutation tests and additional chi-square tests are used \cite{kangadditional}
These statistical tests are conducted under the null hypothesis that the
given sample sequence of the PRNG?s output is independent
and identically distributed with the uniform distribution.
 \cite{kangadditional}


Hamano and Kaneko [6] have shown that the
overlapping template matching test included in the NIST SP
800-22 randomness test suite uses the inaccurate occurrence
probabilities of the template.
 \cite{kangadditional} direkete Quelle bereits bekannt
In this paper, we concentrate on the additional chi-square tests and analyze them from the view point of the nonparametric
 statistical method \cite{kangadditional}
The experimental results show that the additional chi-square tests with the corrected degrees of freedom are more effective. \cite{kangadditional}
 Since we cannot obtain an appopriate hypothesis about the distribution of the noise source which is a component of the entropy source model, assume the distribution of the entropy source is unknown. \cite{kangadditional}




\section{Discussion}

'It has been shown that the overlapping template matching test included in the NIST randomness test suite uses the inaccurate occurrence probabilities PI of the template.'
\cite{hamano2007correction,chen2016templates}. \cite{chen2016templates} proposal of approach, identification of relevant templates by correlation check. If templates B \&  B' have low correlation
both are selected for a test. If the correlation is high just one of both is chosen, as it is assumed to represent the discarded to the greatest extent. This approach reduces the amount of templates significantly (50 \%) \cite{chen2016templates}.

------------

BIOS serial number is ignored

------------

cycles and jiffies may provide more entropy at ungleichermassiger CPU load

------------

interrupt analyse ocnfidence lvele 0.05 ausreichend ?

------------
zu geringe timing variazen die dazu fuehren dass hardware vents nicht verabreitet werden bleiben unberuecksichtigt
------------

 Wenn mehr samples entnommen als bis KEETYPERANDOMINTECRETSET eintritt (wegen Number od reuiqred Samples) darauf hinweisen.
 
 ------------------------
 nicht nachvollziehbar:
  The entropy sources use emulated hardware in the guest. In cloud systems, the only source of entropy(100%) is disk \cite{kumari2015analysis}
 @phdthesis{kumari2015analysis,
     title={Analysis of Linux Random Number Generator in Virtualized Environment},
     author={Kumari, Rashmi},
     year={2015},
     school={University of Calgary}
 }
 
